Improving credibility and efficiency of decision-making \\ within democratic communities via a statistical quorum

### 2018-04-19 statistical-quorum.tex
Adding data.

The rest of the paper is structured as follows. In Section~\ref{sec:related-work} we
present related work. We continue by describing a framework of computing statistical
credibility of results and statistical quorum in Section~\ref{sec:decision-credibility}.
We present our methodology in Section~\ref{sec:methodology} after which we do qualitative and
quantitative evaluation in Section~\ref{sec:evaluation}. We discuss results in
Section~\ref{sec:discussion} and conclude in Section~\ref{sec:conclusion}. At the end we
have Appendix where we derive formulas presented in Section~\ref{sec:decision-credibility}.

@@

PeerMind source code is available under open source licence at \url{https://redacted.for.blind.review} and
anonymized dataset of interactions with the app, including votes, is available at \url{https://redacted.for.blind.review}.

@@

We evaluate statistical quorum in two ways. First, we analyze it theoretically and explore its mathematical properties.
Then we do qualitative and quantitative evaluation through real-world use of the PeerMind application by a 140 members
community.

@@

@@

We can see that in fall members were more satisfied with the decision making in the community and only how understandable
is the whole process decreased.
We attribute this to the fact that the use of the web application is
a novelty, including using a novel concept like statistical quroum, and requires more effort
in educating members than a traditional decision making.
Moreover, decision making in person allows one to observe others how they are participating and
learn from them, instead of being alone using the application.
We explore this question also through interviews in Section~\ref{sec:interviews}.

@@

>>>>

The rest of the paper is structured as follows. In Section~\ref{sec:related-work} we
present related work. We continue by describing a framework of computing statistical
credibility of results and statistical quorum in Section~\ref{sec:decision-credibility}.
We present our methodology in Section~\ref{sec:methodology} after which we do
quantitative and qualitative evaluation in Section~\ref{sec:evaluation}. We discuss results in
Section~\ref{sec:discussion} and conclude in Section~\ref{sec:conclusion}. At the end we
have Appendix where we derive formulas presented in Section~\ref{sec:decision-credibility}.

@@

PeerMind source code is available under open source licence at \url{https://redacted.for.blind.review} and
anonymized dataset of interactions with the app, including votes, is available at \url{https://redacted.for.blind.review}.
The research protocol with human subjects was approved by \emph{redacted for blind review} under
ID \emph{redacted for blind review}.

@@

We evaluate statistical quorum in two ways. First, we analyze it theoretically and explore its mathematical properties.
Then we do quantitative and qualitative evaluation through real-world use of the PeerMind application by a 140 members
community.

@@

We tracked web application activity of 132 PeerMind users who consented to the experiment.
We looked at time they had the web application focused in their browsers to determine the
time they used the web application.
In total, they spent 924.886 seconds using the web application (which translates to roughly 2 hours
per member on average) during the fall 2016 semester.

In comparison, we analyzed spring 2016 council minutes when community was meeting in-person and
estimated that members spent approximately 2.600.000 seconds total doing decision making
(around 5 hours per member on average).
We estimated this based on information about beginning and end of councils multiplied
with an estimation of the number of people attending each council.
We estimated the number of people attending based on number of votes made for different proposals
and the quorum size, taking the maximum of both.
We find this a good and conservative proxy because councils have a rule that they have to adjurn
if the number of members present fall under the quorum (42 members) for more than 10 minutes
which in practice members discover soon through the number of votes being made or when count
is explictly requested.

Users on PeerMind made 139 proposals over the span of a semester (this figure
excludes proposals that were withdrawn before voting began).
Out of those 139 proposals, 98 proposals passed. In comparison, 70 out of 73 proposals passed
in the semester prior.

PeerMind users made 3958 votes, whereas 1396 votes were made during spring semester.
Even after adjusting for the larger number of proposals made in the fall, participation
increased significantly.
From around 19 votes per proposal to 28 votes per proposal, on average.

Plotting time spent per user on PeerMind gives further insights into the democratization of decision-making. While the average person spent 7002 seconds on the application, the distribution of users is fairly long-tailed with many people only spending 30 minutes and some people spending upwards of 20 hours. We believe many people who only spent 30 minutes would not participate in the offline council at all (which is confirmed by much lower attendance and fewer votes). Furthermore, people who need more time to discuss are given the freedom to do so without wasting time of those who are not interested in the pertinent discussion.

@@

We can see that in fall (using PeerMind) members were more satisfied with the decision making in the community and only how understandable
is the whole process decreased.

@@

decline in understanding of the process:

We attribute this to the fact that the use of the web application is
a novelty, including using a novel concept like statistical quorum, and requires more effort
in educating members than a traditional decision making.
Moreover, decision making in person allows one to observe others how they are participating and
learn from them, instead of being alone using the application.
We explore this question also through interviews in Section~\ref{sec:interviews}.

This not only means that PeerMind users were more productive in terms of decisions made, they also spent significantly less time per motion than the control group.
